---
title: "Load data easily and efficiently with lissyuse"
author: "Gonçalo Marques - Luxembourg Income Study"
format: html
editor: visual
---

## Description

The function `lissyuse()` is the newest tool in the `lissyrtools` package, designed to help LISSY R-users quickly import entire datasets from a specific country and/or time period.

It also allows user to specify which variables to import, along with a set of default variables (IDs, weights, currency, year, relation, etc.). If both household-level and person-level variables are specified, `lissyuse()` will automatically merge the two types of files. For faster and more efficient processing, we strongly recommend selecting of a restricted set of variables in `vars` argument. Additionally, the function includes a `subset` argument that allows users to limit the data to a specific subgroup.

Please refer to the following links on [how to start with LISSY](https://www.lisdatacenter.org/data-access/lissy/), and on LIS's metadata information system: [METIS](https://www.lisdatacenter.org/frontend#/home) for detailed information on our available datasets and their respective variables.

## Usage

```{r, eval = FALSE}
lissyuse(
  data = NULL, 
  vars = NULL, 
  from = NULL, 
  to = NULL, 
  subset = NULL, 
  lws = FALSE
)  
```

## Arguments

+----------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `data`   | A character vector containing the datasets to be loaded.                                                                                                                                                                                                    |
|          |                                                                                                                                                                                                                                                             |
|          | 1.  One can select the datasets separately, using their *ccyy* format, e.g., `data = c("de16", "it20")`; \| \|                                                                                                                                              |
|          |                                                                                                                                                                                                                                                             |
|          | 2.  To select and entire series from a given country, we recommend the use of the format *cc*, e.g., `data = c("us", "uk")`; \| \| \|                                                                                                                       |
|          |                                                                                                                                                                                                                                                             |
|          | 3.  These two formats can also be used simultaneously, e.g., `data = c("be", "es", "br21", "br20")`**.** \| \|                                                                                                                                              |
+----------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `vars`   | A character vector specifying the LIS/LWS variables to be loaded, e.g., `vars = c("dhi", "region_c", "age", "relation")`.                                                                                                                                   |
|          |                                                                                                                                                                                                                                                             |
|          | If no variables are specified, the entire dataset will be loaded.                                                                                                                                                                                           |
|          |                                                                                                                                                                                                                                                             |
|          | If only household-level variables are selected (e.g., `c("dhi", "region_c")`), the data will be loaded at the household level. However, if at least one person-level variable is included (e.g., `"age"`), the data will be loaded at the individual level. |
|          |                                                                                                                                                                                                                                                             |
|          | When selecting specific variables, note that some additional variables will be included by default. For clarity, run `lissyrtools::key_vars_household_lis`, `lissyrtools::key_vars_person_lis`, and their equivalents for LWS.                              |
+----------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `from`   | A numeric value representing the year (inclusive) after which the LIS/LWS datasets should be loaded.                                                                                                                                                        |
+----------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `to`     | A numeric value representing the year (inclusive) up to which the LIS/LWS datasets should be loaded.                                                                                                                                                        |
+----------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `subset` | A logical expression defining the criteria for subsetting the data. Observations for which the expression evaluates to `TRUE` are included in the subset.                                                                                                   |
+----------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| `lws`    | A logical value indicating whether to load LWS data. If `TRUE`, LWS data is loaded; otherwise (default: `FALSE`), LIS data is loaded instead. Note that this does not eliminate the need to set the 'Project' field accordingly in the LISSY remote system. |
+----------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

## Value

A list named `lis_datasets` or `lws_datasets`. Each element of the list will be a data frame named after their respective dataset. See the naming formats in the examples below. Each data frame will contain as many columns as the selected variables, plus the default technical ones.

## Examples

```{r, eval = FALSE}
{ech}

library(lissyrtools)
library(dplyr)

# ----------------------------- LIS --------------------------------------- 

# Loading a list of data frames of using household-level and person-level variables 

lissyuse( 
  data = c("it", "de16", "us19"), 
  vars  = c("dhi", "region_c", "age", "hourstot", "status1"), 
  subset = "!is.na(status1) & relation %in% c(1000,2000)"
)

# Checking the names of the data frames. 
names(lis_datasets)

# Selecting certain elemennts of the list 
lis_datasets[["it14"]]   # By their name
lis_datasets[1:3]        # By their respective order within lis_datasets

# Selecting all the italian datasets, while restrict them to a certain year range. 
lissyuse(
  data = c("it"), 
  vars  = c("dhi", "region_c"), 
  from = 2004, 
  to = 2016
)

# In the previous line only household-level variables were selected. 
  # this will lead to slightly different names for the data frames. 
names(lis_datasets)

# The same occurs when only person-level variables were selected 
lissyuse(
  data = c("it"), 
  vars  = c("age", "sex"), 
  from = 2004, 
  to = 2016
)

names(lis_datasets)




# ----------------------------- LWS --------------------------------------- 

# Example for LWS 
lissyuse(
  data = c("us", "uk17", "uk19"), 
  vars = "dnw", 
  from = 2015, 
  to = 2021,
  lws = TRUE
)

# When working with LWS datasets, the list will be named lws_datasets
names(lws_datasets)
 

```
